---
title: "Introduction to  CAPPMx"
author:
  - Noirrit Kiran Chandra^[noirritchandra@gmail.com]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to  CAPPMx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Load the package:

```{r setup}
library(CAPPMx)
```

## Illustration with both Categorical and Continuous Covariates

#### Simulation Setup

We set the sample size and dimension.

```{r set dimensions}
p=10 #total number of covariates
p_cat=3; # number of categorical covariates
p_cont=p-p_cat # number of continuous covariates
nsamp1=100 #number of subjects in the Trt arm 
samp.ratio=6 
nsamp2=samp.ratio * nsamp1 #number of subjects in the RWD arm 
```

We set the sample size in the RWD considerably larger than that in the treatment arm. Then we generate the covariates from mixture model. We consider a 2-component mixture in the treatment arm and a 3-component mixture in RWD such that the atoms of the mixtures in the RWD form a superset of the atoms of the mixtures in the treatment arm.

```{r message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
set.seed(1)
library(mvtnorm)
k2=2 ####number of true cluster -1

##############sample mixture assignments########
lab1= sample.int(n=k2,size=nsamp1, replace = T) #
lab2= sample(1:(k2+1),size=nsamp2, replace = T,prob = c(1,1,4))
################################################

##set atoms for the mixtures of the coninuous covariates
mu=matrix(0,nrow=k2+1,ncol=p-3)
mu[1,1:2]=mu[2,5:6]=2
########################################################

######Binary categorical covariates
prob_trt=.6 
prob_rwd=.85
#########################

X1.cont=matrix(0, nrow=nsamp1,ncol=p_cont)
X2.cont=matrix(0, nrow=nsamp2,ncol=p_cont); X2.cat=matrix(0, nrow=nsamp2,ncol=p_cat);

var_each_mix=.05
##########Generate from case (iii) Lamb model ########################
# lambda=simulate_lambda(d,p,1)
for(i in 1:(k2+1)){
  ####set RWD
  inds=which(lab2==i)
  if(length(inds>0)){
    X2.cont[inds,]= rmvnorm(length(inds),mean = mu[i,],sigma = var_each_mix*diag(p_cont ) )
    if(i<=k2) ###set categorical covs in RWD 
      X2.cat[inds,]= matrix(rbinom(p_cat*length(inds), size=1,prob = prob_trt),
            nrow=length(inds)) else X2.cat[inds,]= matrix(rbinom(p_cat*length(inds),
          size=1,prob = prob_rwd)  ,nrow=length(inds))
  }
  rm(inds)
  ###########
  ####set Trt Arm
  inds=which(lab1==i)
  if(length(inds>0))
    X1.cont[inds,]= rmvnorm(length(inds),mean = mu[i,],sigma = var_each_mix*diag(p_cont ) )
  rm(inds)
  ###########
  # print(i)
}

###set categorical covs in trt arm
X1.cat=matrix(rbinom(p_cat*nsamp1, size=1,prob = prob_trt)  ,nrow=nsamp1)
```

Thus we have generated the following:

-   `X1.cont` - continuous covariates in the treatment arm.

-   `X1.cat` - categorical covariates in the treatment arm.

-   `X2.cont` - continuous covariates in the RWD.

-   `X2.cat` - categorical covariates in the RWD.

Next we sample responses using a linear model

```{r}
X1=cbind(X1.cont, X1.cat)
X2=cbind(X2.cont, X2.cat)
# rm(X1.cont, X1.cat, X2.cont,X2.cat)
#########################################

delta=1 ##treatment effect

#sample true regression coefficients for the linear response model
beta= sample(c(-1,2),size=p,replace = T ) 
beta[sample.int(p,5)]=0

y1=rnorm(X1%*%beta)+delta #response in the treatment arm
y2=rnorm(X2%*%beta) #response in the RWD
rm(X1,X2)
```

We deliberately make 5% entries missing in each covariate vector

```{r}
make_missing=function(x, missing_percent=.05){
  x.mis=x
  x.mis[sample.int(n=length(x),size=floor(length(x)*.05)) ]=NA
  x.mis
}

X1.cat.mis=make_missing(X1.cat);X1.cont.mis=make_missing(X1.cont)
X2.cat.mis=make_missing(X2.cat);X2.cont.mis=make_missing(X2.cont)
```

#### CA-PPMx Fitting

Then we fit the CA-PPMx model

```{r fit_cappmx,message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
result=cappmx_fit(X1.cat.mis, #categorical covariates in the trt arm
        X1.cont.mis, #continuous covariates in the trt arm
        y1, #observed responses in the trt arm
        rep(TRUE,length(y1)), #survival indicators in the trt arm
        X2.cat.mis, #categorical covariates in the RWD
        X2.cont.mis, #continuous covariates in the RWD
        y2, # observed responses in the RWD
        rep(TRUE,length(y2)) #survival indicators in the RWD
        )
```

#### Investigate Population Equivalence

We compare the area under the curve (AUC) values using Bayesian additive regression tree (BART) classifiers for our *importance-resampling adjusted* synthetic control population versus *randomly resampled* population.

```{r find_auc, message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
aucs=get_auc(cat_cov_trt=X1.cat.mis,cont_cov_trt=X1.cont.mis,
                 cat_cov_rwd=X2.cat.mis,cont_cov_rwd=X2.cont.mis, result)
```

```{r show_aucs}
aucs
```

Next we proceed towards model-based analysis on treatment effects.

#### Bayesian Model-Based Inference on Treatment Effect

Estimate the *average treatment effect* from the MCMC outputs stored in `result`

```{r}
average_trt_effect(result)
```

## Illustration with *ONLY Categorical* Covariates

We use the previously sampled categorical covariates. Then we generate the response

```{r }
X1=cbind( X1.cat)
X2=cbind(X2.cat)
#########################################

delta=1 ##treatment effect

#sample true regression coefficients for the linear response model
beta= c(-1,0,2) 

y1=rnorm(X1%*%beta)+delta #response in the treatment arm
y2=rnorm(X2%*%beta) #response in the RWD
rm(X1,X2)
```

#### CA-PPMx Fitting

Then we fit the CA-PPMx model

```{r fit_cappmx_cat_only,message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
result=cappmx_fit(cat_cov_trt=X1.cat.mis, #categorical covariates in the trt arm
         cont_cov_trt= NULL, #we set the continuous covariate=NULL
        y1, #observed responses in the trt arm
        rep(TRUE,length(y1)), #survival indicators in the trt arm
        cat_cov_rwd=X2.cat.mis, #categorical covariates in the RWD
        cont_cov_rwd=NULL, #continuous covariates in the RWD
        y2, # observed responses in the RWD
        rep(TRUE,length(y2)) #survival indicators in the RWD
        )
```

#### Investigate Population Equivalence

We compare the area under the curve (AUC) values using Bayesian additive regression tree (BART) classifiers for our *importance-resampling adjusted* synthetic control population versus *randomly resampled* population.

```{r find_auc_cat_only, message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
aucs=get_auc(cat_cov_trt=X1.cat.mis, #categorical covariates in the trt arm
        cat_cov_rwd=X2.cat.mis, #categorical covariates in the RWD
        result=result)
```

```{r show_aucs_cat_only}
aucs
```

Next we proceed towards model-based analysis on treatment effects.

#### Bayesian Model-Based Inference on Treatment Effect

Estimate the *average treatment effect* from the MCMC outputs stored in `result`

```{r}
average_trt_effect(result)
```

## Illustration with *ONLY Continuous* Covariates

We use the previously sampled categorical covariates. Then we generate the response

```{r }
X1=cbind( X1.cont)
X2=cbind(X2.cont)
#########################################

delta=1 ##treatment effect

#sample true regression coefficients for the linear response model
beta= sample(x=c(-1,0,2),size=ncol(X1),replace = T,prob = c(1,2,1)) 

y1=rnorm(X1%*%beta)+delta #response in the treatment arm
y2=rnorm(X2%*%beta) #response in the RWD
rm(X1,X2)
```

#### CA-PPMx Fitting

Then we fit the CA-PPMx model

```{r fit_cappmx_cont_only,message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
result=cappmx_fit(cont_cov_trt= X1.cont.mis, 
         response_trt=y1, #observed responses in the trt arm
        surv_ind_trt=rep(TRUE,length(y1)), #survival indicators in the trt arm
        cont_cov_rwd=X2.cont.mis, #continuous covariates in the RWD
        response_rwd=y2, # observed responses in the RWD
        surv_ind_rwd=rep(TRUE,length(y2)) #survival indicators in the RWD
        )
```

#### Investigate Population Equivalence

We compare the area under the curve (AUC) values using Bayesian additive regression tree (BART) classifiers for our *importance-resampling adjusted* synthetic control population versus *randomly resampled* population.

```{r find_auc_cont_only, message=FALSE, results=FALSE, warning=FALSE, comment=FALSE}
aucs=get_auc(cont_cov_trt= X1.cont.mis, 
        cont_cov_rwd=X2.cont.mis, #continuous covariates in the RWD,
        result=result)
```

```{r show_aucs_cont_only}
aucs
```

Next we proceed towards model-based analysis on treatment effects.

#### Bayesian Model-Based Inference on Treatment Effect

Estimate the *average treatment effect* from the MCMC outputs stored in `result`

```{r}
average_trt_effect(result)
```

